---
title: "README"
output:
  github_document:
    toc: true
---

```{r include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
```

# happyCompare

Reporting toolbox for happy output



## Demo datasets

#### `happyCompare_list`

A demo `happyCompare_list` object, obtained by running hap.py on hap.py example data and loading the results into R using `happyCompare::read_samplesheet()`.

**Step 1: run hap.py**

```
$HAPPY_ROOT=/path/to/hap.py/repo/root
$HAPPY_BIN=/path/to/hap.py/binnary

# fixed hap.py resources
truth=$HAPPY_ROOT/example/happy/PG_NA12878_hg38-chr21.vcf.gz
conf=$HAPPY_ROOT/example/happy/PG_Conf_hg38-chr21.bed.gz
ref=$HAPPY_ROOT/example/happy/hg38.chr21.fa

# query VCFs
gatk_q=$HAPPY_ROOT/example/happy/NA12878-GATK3-chr21.vcf.gz
freebayes_q=$HAPPY_ROOT/example/happy/NA12878-Freebayes-chr21.vcf.gz
platypus_q=$HAPPY_ROOT/example/happy/NA12878-Platypus-chr21.vcf.gz

# run hap.py
mkdir -p $outdir
$HAPPY_BIN -f $conf -r $ref --no-json $truth $gatk_q --roc-filter LowQual --roc QUAL -o $outdir/gatk
$HAPPY_BIN -f $conf -r $ref --no-json $truth $freebayes_q --roc-filter LowQual --roc QUAL -o $outdir/freebayes
$HAPPY_BIN -f $conf -r $ref --no-json $truth $platypus_q --roc QUAL -o $outdir/platypus
```

**Step 2: Create a happyCompare samplesheet**

```
Group.Id,Sample.Id,Replicate.Id,happy_prefix
freebayes,NA12878,NA12878_freebayes,output/freebayes
gatk,NA12878,NA12878_gatk,output/gatk
platypus,NA12878,NA12878_platypus,output/platypus
```

**Step 3: Load data into R**

```{r eval=FALSE}
library(happyCompare)

happyCompare_list = read_samplesheet(samplesheet_path = "happyCompare/happyCompare_samplesheet.csv")
save(happyCompare_list, file = "happyCompare/happyCompare_list.RData")
```

## Examples

### Static reports

**Example germline report:** [germline.html](inst/examples/germline.html)

```{bash, eval=TRUE, error=FALSE, results='hide'}
DATA="C:/Users/mgonzalez/SublimeProjects/happy-example-data"
HAPPYCOMPARE="C:/Users/mgonzalez/SublimeProjects/happyCompare"
RESULTS_ROOT="C:/Users/mgonzalez/SublimeProjects/happy-example-data"
Rscript --vanilla $HAPPYCOMPARE/exec/happyCompare.R \
        --input_template $HAPPYCOMPARE/inst/rmd/germline.Rmd \
        --output_dir $HAPPYCOMPARE/inst/examples \
        --samplesheet $DATA/happyCompare/happyCompare_samplesheet.csv \
        --root_dir $RESULTS_ROOT
```

```
## [DONE] Output: C:/Users/mgonzalez/SublimeProjects/happyCompare/inst/examples
```