---
title: "Analysing stratified counts"
author: "Mar Gonzalez-Porta"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calibrating HDIs with simulated data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.width = 6, cache = FALSE)

library("happyCompare")
library("tidyverse")
library("ggplot2")
theme_set(theme_bw())
```

## Set up

In the present vignette we will be comparing stratified recall between a set of PCR-Free vs. Nano builds from NA12878. Starting from publicly available VCFs in [BaseSpace Data Central](https://basespace.illumina.com/datacentral), we can obtain stratified counts with hap.py as follows:

```
hap.py ${TRUTH_VCF} ${QUERY_VCF} -o ${OUTPUT_PREFIX} -f ${CONFIDENT_REGIONS} \
    --threads 40 --write-counts --stratification filtered_beds/stratification_config.txt --reference ${REF} \
    --roc QUAL --roc-filter LowQual --no-json
```

Then load the output to R using `happyCompare`:

```{r message=FALSE}
samplesheet_path <- system.file("extdata/samplesheets", "pcrfree_vs_nano.vignettes.csv", package = "happyCompare")
happy_compare <- read_samplesheet(samplesheet_path)
```

We will eventually end up with a `happy_compare` object (a list), with 3 elements: 
```{r}
class(happy_compare)
sapply(happy_compare, class)
```

If we furhter inspect the contents of `happy_results`, we can see that each of its elements matches the data structures described in `happyR`, e.g.:

```{r}
names(happy_compare$happy_results[[1]])
```


## Extracting data from `happy_compare` objects

Once we have a `happy_compare` object we can rely on the `extract_metrics` function to access our metrics of interest. Since hap.py saves stratified counts under `*extended.csv` files, we can use the following:

```{r }
stratified_counts <- extract_metrics(happy_compare, table = "extended") %>% 
  filter(Subtype == "*", Filter == "PASS", Subset.Level == 0, !grepl(pattern = "TS*", Subset)) 
```

This will conveniently merge hap.py results with samplesheet metadata, making downstream analysis easier:
```{r}
stratified_counts
```

## Downstream analysis of stratified counts

Finally, we can estimate highest density intervals to account for uncertainty in the stratified counts:

```{r message=FALSE}
hdi <- stratified_counts %>% 
  estimate_hdi(successes_col = "TRUTH.TP", totals_col = "TRUTH.TOTAL", 
               group_cols = c("Group.Id", "Subset", "Type"), aggregate_only = FALSE)
```

And visualise the results using `ggplot2`:

```{r}
hdi %>% 
  mutate(Subset = factor(Subset, levels = rev(unique(Subset)))) %>% 
  filter(replicate_id == ".aggregate") %>% 
  ggplot(aes(x = estimated_p, y = Subset, group = Subset)) +
    geom_point(aes(color = Group.Id), size = 2) +
    geom_errorbarh(aes(xmin = lower, xmax = upper, color = Group.Id), height = 0.4) +
    facet_grid(. ~ Type) +
    theme(legend.position = "bottom") +
    ggtitle("Recall estimates across L0 subsets") +
    ylab("")
```